Beliefs about the development of mental life

### 2025-01-06 paper/paper.Rmd
typos and revisions

Compared to these attributions of _cognition and control_, participants attributed substantially greater capacities to newborns in the domains that we have labeled _social connection_ (median score for newborns: `r s2_cap_rat_newborns_ss %>% filter(factor == "Social connection") %>% pull(median) %>% round_x(., 2)` out of 100 points, $M$ = `r s2_cap_rat_newborns_ss %>% filter(factor == "Social connection") %>% pull(mean) %>% round_x(., 2)`, $sd$ = `r s2_cap_rat_newborns_ss %>% filter(factor == "Social connection") %>% pull(sd) %>% round_x(., 2)`; increase in log-odds relative to _cognition and control_: $\beta$ = `r round_x(s2_gam_ptab["factorsoc_cog", "Estimate"], 2)`, $p$ `r ifelse(s2_gam_ptab["factorsoc_cog", "Pr(>|z|)"] < 0.001, " < 0.001", paste(" = ", round_x(s2_gam_ptab["factorsoc_cog", "Pr(>|z|)"], 3)))`), but still perceived dramatic non-linear growth in this domain over the target age range ($\chi^2$ = `r s2_gam_parameters %>% filter(Parameter == "s(target_year):factorSocial connection") %>% pull(z) %>% round_x(., 2)`, $p$ `r ifelse(s2_gam_parameters %>% filter(Parameter == "s(target_year):factorSocial connection") %>% pull(p)  < 0.001, " < 0.001", paste(" = ", round_x(s2_gam_parameters %>% filter(Parameter == "s(target_year):factorSocial connection") %>% pull(p), 3)))`), $edf$ = `r s2_gam_parameters %>% filter(Parameter == "s(target_year):factorSocial connection") %>% pull(df) %>% round_x(., 2)`), characterized by rapid development over the child's first year of life (see Figure 2, Panel A, third box from left).

@@

Because Study 3 included additional questions probing participants' theories of developmental mechanisms, we reduced the set of mental capacities from 20 in Study 2 to eight (two capacities per domain) in the current study to maintain an appropriate level of participant burden. We chose "controlling their emotions" and "reasoning about things" to represent the domain of _cognition and control_; "getting hungry" and "feeling pain" to represent the domain of _bodily sensation_; "learning from other people" and "feeling happy" to represent the domain that we have labeled _social connection_; and "feeling distressed" and "feeling helpless" to represent the domain that we have labeled _negative affect_. Capacities were presented in a random order for each participant.

@@

This approach yielded a very similar picture of US adults' understanding of the development of human mental life; see Figure 2, Panel B. As in Study 2, participants perceived that newborns have very little capacity for _cognition and control_ (median score for newborns: `r s3_cap_rat_newborns_ss %>% filter(factor == "Cognition and control") %>% pull(median) %>% round_x(., 2)` out of 100 points, $M$ = `r s3_cap_rat_newborns_ss %>% filter(factor == "Cognition and control") %>% pull(mean) %>% round_x(., 2)`, $sd$ = `r s3_cap_rat_newborns_ss %>% filter(factor == "Cognition and control") %>% pull(sd) %>% round_x(., 2)`), but reported that such capacities increase dramatically over the target age range ($\chi^2$ = `r s3_gam_parameters %>% filter(Parameter == "s(target_year):factorCognition and control") %>% pull(z) %>% round_x(., 2)`, $p$ `r ifelse(s3_gam_parameters %>% filter(Parameter == "s(target_year):factorCognition and control") %>% pull(p)  < 0.001, " < 0.001", paste(" = ", round_x(s3_gam_parameters %>% filter(Parameter == "s(target_year):factorCognition and control") %>% pull(p), 3)))`). Again, participants attributed substantially greater capacities to newborns in the domains that we have labeled _social connection_ (median score for newborns: `r s3_cap_rat_newborns_ss %>% filter(factor == "Social connection") %>% pull(median) %>% round_x(., 2)` out of 100 points, $M$ = `r s3_cap_rat_newborns_ss %>% filter(factor == "Social connection") %>% pull(mean) %>% round_x(., 2)`, $sd$ = `r s3_cap_rat_newborns_ss %>% filter(factor == "Social connection") %>% pull(sd) %>% round_x(., 2)`; increase in log-odds: $\beta$ = `r round_x(s3_gam_ptab["factorsoc_cog", "Estimate"], 2)`, $p$ `r ifelse(s3_gam_ptab["factorsoc_cog", "Pr(>|z|)"] < 0.001, " < 0.001", paste(" = ", round_x(s3_gam_ptab["factorsoc_cog", "Pr(>|z|)"], 3)))`) and _negative affect_ (median score for newborns: `r s3_cap_rat_newborns_ss %>% filter(factor == "Negative affect") %>% pull(median) %>% round_x(., 2)` out of 100 points, $M$ = `r s3_cap_rat_newborns_ss %>% filter(factor == "Negative affect") %>% pull(mean) %>% round_x(., 2)`, $sd$ = `r s3_cap_rat_newborns_ss %>% filter(factor == "Negative affect") %>% pull(sd) %>% round_x(., 2)`; increase in log-odds: $\beta$ = `r round_x(s3_gam_ptab["factorneg_cog", "Estimate"], 2)`, $p$ `r ifelse(s3_gam_ptab["factorneg_cog", "Pr(>|z|)"] < 0.001, " < 0.001", paste(" = ", round_x(s3_gam_ptab["factorneg_cog", "Pr(>|z|)"], 3)))`); in Study 3, as in Study 2, participants reported substantial increases in the domain that we have labeled _social connection_ ($\chi^2$ = `r s3_gam_parameters %>% filter(Parameter == "s(target_year):factorSocial connection") %>% pull(z) %>% round_x(., 2)`, $p$ `r ifelse(s3_gam_parameters %>% filter(Parameter == "s(target_year):factorSocial connection") %>% pull(p)  < 0.001, " < 0.001", paste(" = ", round_x(s3_gam_parameters %>% filter(Parameter == "s(target_year):factorSocial connection") %>% pull(p), 3)))`) and relatively small increases over the target age range in the domain of _negative affect_ ($\chi^2$ = `r s3_gam_parameters %>% filter(Parameter == "s(target_year):factorNegative affect") %>% pull(z) %>% round_x(., 2)`, $p$ `r ifelse(s3_gam_parameters %>% filter(Parameter == "s(target_year):factorNegative affect") %>% pull(p)  < 0.001, " < 0.001", paste(" = ", round_x(s3_gam_parameters %>% filter(Parameter == "s(target_year):factorNegative affect") %>% pull(p), 3)))`). Finally, as in Study 2, in the domain of _bodily sensation_ participants attributed nearly maximal capacity to newborns (median score for newborns: `r s3_cap_rat_newborns_ss %>% filter(factor == "Bodily sensation") %>% pull(median) %>% round_x(., 2)` out of 100 points, $M$ = `r s3_cap_rat_newborns_ss %>% filter(factor == "Bodily sensation") %>% pull(mean) %>% round_x(., 2)`, $sd$ = `r s3_cap_rat_newborns_ss %>% filter(factor == "Bodily sensation") %>% pull(sd) %>% round_x(., 2)`; increase in log-odds relative to _Social connection_: $\beta$ = `r round_x(s3_gam_ptab["factorbod_cog", "Estimate"], 2)`, $p$ `r ifelse(s3_gam_ptab["factorbod_cog", "Pr(>|z|)"] < 0.001, " < 0.001", paste(" = ", round_x(s3_gam_ptab["factorbod_cog", "Pr(>|z|)"], 3)))`) and small increases in capacities for _bodily sensation_ over the target age range ($\chi^2$ = `r s3_gam_parameters %>% filter(Parameter == "s(target_year):factorBodily sensation") %>% pull(z) %>% round_x(., 2)`, $p$ `r ifelse(s3_gam_parameters %>% filter(Parameter == "s(target_year):factorBodily sensation") %>% pull(p)  < 0.001, " < 0.001", paste(" = ", round_x(s3_gam_parameters %>% filter(Parameter == "s(target_year):factorBodily sensation") %>% pull(p), 3)))`).

Again, we note that this modeling approach did not include formal statistical tests of differences in perceived trajectories across domains—but visual inspection of Figure 2, Panel B, makes it clear that the shape of these perceived developmental trajectories varied substantially across domains, echoing quite precisely the results of our post-hoc exploratory analyses in Study 2.

@@

Why do US adults believe certain aspects of mental life, but not others, to be present at birth? What do US adults perceive to be driving the development of capacities for what we have called _bodily sensation_, _negative affect_, _social connection_, and _cognition and control_? As a first step in what we hope to be a longer line of work addressing such questions, we asked participants to assess the importance of a variety of possible developmental mechanisms for each of the capacities included in this study. We theorized that when people perceive that a capacity is substantially present at birth, they believe that the capacity is innate, and, conversely, that when people perceive that a capacity develops slowly over years, they believe the capacity is in some sense learned. [For in-depth investigations of laypeople's intuitions about the innateness of cognitive abilities, motor abilities, and emotions, see @wang.feigenson2019; @berent.etal2019; @berent.etal2020.]

@@

Meanwhile, the explanation that "the child's brain changes (for example, brain grows bigger, more or fewer connections between neurons)" was a modal choice for `r s3_mech_choice_modes %>% filter(dev_factor == "brain_changes") %>% nrow()` of our 8 capacities, including helplessness (in the domain of _negative affect_) and social learning (_social connection_), as well as emotional control and reasoning things (both in the domain of _cognition and control_). However, consensus about this selection was much lower (`r (s3_mech_choice_modes %>% filter(dev_factor == "brain_changes") %>% top_n(-1, prop) %>% pull(prop) %>% round(., 2))*100`-`r (s3_mech_choice_modes %>% filter(dev_factor == "brain_changes") %>% top_n(1, prop) %>% pull(prop) %>% round(., 2))*100`% of participants across capacities). Instead, these capacities were characterized by a more diverse selection of mechanisms, including mechanisms that are not as directly tied to biology: For example, when considering what drives the development of a capacity to learn from other people, roughly equal numbers of participants selected brain changes (`r (s3_mech_choice_modes %>% filter(dev_factor == "brain_changes", grepl("learn", capacity)) %>% pull(prop) %>% round(., 2))*100`%), observations of the social world (`r (s3_mech_choice_ss_c %>% filter(grepl("observes people", dev_factor_text), grepl("learn", capacity)) %>% pull(prop) %>% round(., 2))*100`%), and interactions with others (`r (s3_mech_choice_ss_c %>% filter(grepl("interacts", dev_factor_text), grepl("learn", capacity)) %>% pull(prop) %>% round(., 2))*100`%); and when considering the capacity for controlling one's emotions, participants were nearly as likely to select explicit teaching (`r (s3_mech_choice_ss_c %>% filter(grepl("teach", dev_factor_text), grepl("control", capacity)) %>% pull(prop) %>% round(., 2))*100`%) as the most important developmental mechanism as they were to select brain changes (`r (s3_mech_choice_modes %>% filter(dev_factor == "brain_changes", grepl("control", capacity)) %>% pull(prop) %>% round(., 2))*100`%).

@@

For capacities in the domain that we have called _bodily sensation_ (Figure 3, top row), "the child is biologically 'preprogrammed' to have this ability" stood out as the highest-rated mechanism by far (median score: `r s3_mech_rat_ss_c %>% top_n(1, mean) %>% pull(median) %>% as.numeric() %>% round_x(., 2)` on a scale from 0-6, $M$ = `r s3_mech_rat_ss_c %>% top_n(1, mean) %>% pull(mean) %>% as.numeric() %>% round_x(., 2)`, sd = `r s3_mech_rat_ss_c %>% top_n(1, mean) %>% pull(sd) %>% as.numeric() %>% round_x(., 2)`; comparison to grand mean: $\beta$ = `r s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("prep", Parameter)) %>% pull(Coefficient) %>% as.numeric() %>% round_x(., 2)`, 95% CI: [`r s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("prep", Parameter)) %>% data.frame() %>% select(CI_low, CI_high) %>% round_x(., 2) %>% unite(CI95, c(CI_low, CI_high), sep = ", ") %>% as.character()`], $p$ `r ifelse(s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("prep", Parameter)) %>% pull(p) < 0.001, " < 0.001", paste(" = ", round_x(s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("prep", Parameter)) %>% pull(p), 3)))`), followed by "the child's brain changes (for example, brain grows bigger, more or fewer connections between neurons)" (median score: `r s3_mech_rat_ss_c %>% top_n(2, mean) %>% top_n(-1, mean) %>% pull(median) %>% as.numeric() %>% round_x(., 2)` on a scale from 0-6, $M$ = `r s3_mech_rat_ss_c %>% top_n(2, mean) %>% top_n(-1, mean) %>% pull(mean) %>% as.numeric() %>% round_x(., 2)`, sd = `r s3_mech_rat_ss_c %>% top_n(2, mean) %>% top_n(-1, mean) %>% pull(sd) %>% as.numeric() %>% round_x(., 2)`; comparison to grand mean: $\beta$ = `r s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("brai", Parameter)) %>% pull(Coefficient) %>% as.numeric() %>% round_x(., 2)`, 95% CI: [`r s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("brai", Parameter)) %>% data.frame() %>% select(CI_low, CI_high) %>% round_x(., 2) %>% unite(CI95, c(CI_low, CI_high), sep = ", ") %>% as.character()`], $p$ `r ifelse(s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("brai", Parameter)) %>% pull(p) < 0.001, " < 0.001", paste(" = ", round_x(s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("brai", Parameter)) %>% pull(p), 3)))`), and "the child's body grows and matures (for example, muscles get stronger, child gets taller)" (median score: `r s3_mech_rat_ss_c %>% top_n(3, mean) %>% top_n(-1, mean) %>% pull(median) %>% as.numeric() %>% round_x(., 2)` on a scale from 0-6, $M$ = `r s3_mech_rat_ss_c %>% top_n(3, mean) %>% top_n(-1, mean) %>% pull(mean) %>% as.numeric() %>% round_x(., 2)`, sd = `r s3_mech_rat_ss_c %>% top_n(3, mean) %>% top_n(-1, mean) %>% pull(sd) %>% as.numeric() %>% round_x(., 2)`; comparison to grand mean: $\beta$ = `r s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("body", Parameter)) %>% pull(Coefficient) %>% as.numeric() %>% round_x(., 2)`, 95% CI: [`r s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("body", Parameter)) %>% data.frame() %>% select(CI_low, CI_high) %>% round_x(., 2) %>% unite(CI95, c(CI_low, CI_high), sep = ", ") %>% as.character()`], $p$ `r ifelse(s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("body", Parameter)) %>% pull(p) < 0.001, " < 0.001", paste(" = ", round_x(s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("body", Parameter)) %>% pull(p), 3)))`). Most of the remaining mechanisms were rated as very unimportant for capacities in this domain, with particularly low ratings for "people explicitly teach the child how to do this" (median score: `r s3_mech_rat_ss_c %>% top_n(-1, mean) %>% pull(median) %>% as.numeric() %>% round_x(., 2)` on a scale from 0-6, $M$ = `r s3_mech_rat_ss_c %>% top_n(-1, mean) %>% pull(mean) %>% as.numeric() %>% round_x(., 2)`, sd = `r s3_mech_rat_ss_c %>% top_n(-1, mean) %>% pull(sd) %>% as.numeric() %>% round_x(., 2)`; comparison to grand mean: $\beta$ = `r s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("teac", Parameter)) %>% pull(Coefficient) %>% as.numeric() %>% round_x(., 2)`, 95% CI: [`r s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("teac", Parameter)) %>% data.frame() %>% select(CI_low, CI_high) %>% round_x(., 2) %>% unite(CI95, c(CI_low, CI_high), sep = ", ") %>% as.character()`], $p$ `r ifelse(s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("teac", Parameter)) %>% pull(p) < 0.001, " < 0.001", paste(" = ", round_x(s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("teac", Parameter)) %>% pull(p), 3)))`).

>>>>

Compared to these attributions of _cognition and control_, participants attributed substantially greater capacities to newborns in the domain that we have labeled _social connection_ (median score for newborns: `r s2_cap_rat_newborns_ss %>% filter(factor == "Social connection") %>% pull(median) %>% round_x(., 2)` out of 100 points, $M$ = `r s2_cap_rat_newborns_ss %>% filter(factor == "Social connection") %>% pull(mean) %>% round_x(., 2)`, $sd$ = `r s2_cap_rat_newborns_ss %>% filter(factor == "Social connection") %>% pull(sd) %>% round_x(., 2)`; increase in log-odds relative to _cognition and control_: $\beta$ = `r round_x(s2_gam_ptab["factorsoc_cog", "Estimate"], 2)`, $p$ `r ifelse(s2_gam_ptab["factorsoc_cog", "Pr(>|z|)"] < 0.001, " < 0.001", paste(" = ", round_x(s2_gam_ptab["factorsoc_cog", "Pr(>|z|)"], 3)))`), but still perceived dramatic non-linear growth in this domain over the target age range ($\chi^2$ = `r s2_gam_parameters %>% filter(Parameter == "s(target_year):factorSocial connection") %>% pull(z) %>% round_x(., 2)`, $p$ `r ifelse(s2_gam_parameters %>% filter(Parameter == "s(target_year):factorSocial connection") %>% pull(p)  < 0.001, " < 0.001", paste(" = ", round_x(s2_gam_parameters %>% filter(Parameter == "s(target_year):factorSocial connection") %>% pull(p), 3)))`), $edf$ = `r s2_gam_parameters %>% filter(Parameter == "s(target_year):factorSocial connection") %>% pull(df) %>% round_x(., 2)`), characterized by rapid development over the child's first year of life (see Figure 2, Panel A, third box from left).

@@

Because Study 3 included additional questions probing participants' theories of developmental mechanisms, to maintain an appropriate level of participant burden we reduced the set of mental capacities from 20 (5 capacities per domain) to 8 (2 capacities per domain). We chose "controlling their emotions" and "reasoning about things" to represent the domain of _cognition and control_; "getting hungry" and "feeling pain" to represent the domain of _bodily sensation_; "learning from other people" and "feeling happy" to represent the domain that we have labeled _social connection_; and "feeling distressed" and "feeling helpless" to represent the domain that we have labeled _negative affect_. Capacities were presented in a random order for each participant.

@@

This approach yielded a very similar picture of US adults' understanding of the development of human mental life; see Figure 2, Panel B. As in Study 2, participants reported that newborns have very little capacity for _cognition and control_ (median score for newborns: `r s3_cap_rat_newborns_ss %>% filter(factor == "Cognition and control") %>% pull(median) %>% round_x(., 2)` out of 100 points, $M$ = `r s3_cap_rat_newborns_ss %>% filter(factor == "Cognition and control") %>% pull(mean) %>% round_x(., 2)`, $sd$ = `r s3_cap_rat_newborns_ss %>% filter(factor == "Cognition and control") %>% pull(sd) %>% round_x(., 2)`), but reported that such capacities increase dramatically over the target age range ($\chi^2$ = `r s3_gam_parameters %>% filter(Parameter == "s(target_year):factorCognition and control") %>% pull(z) %>% round_x(., 2)`, $p$ `r ifelse(s3_gam_parameters %>% filter(Parameter == "s(target_year):factorCognition and control") %>% pull(p)  < 0.001, " < 0.001", paste(" = ", round_x(s3_gam_parameters %>% filter(Parameter == "s(target_year):factorCognition and control") %>% pull(p), 3)))`). Again, participants attributed substantially greater capacities to newborns in the domains that we have labeled _social connection_ (median score for newborns: `r s3_cap_rat_newborns_ss %>% filter(factor == "Social connection") %>% pull(median) %>% round_x(., 2)` out of 100 points, $M$ = `r s3_cap_rat_newborns_ss %>% filter(factor == "Social connection") %>% pull(mean) %>% round_x(., 2)`, $sd$ = `r s3_cap_rat_newborns_ss %>% filter(factor == "Social connection") %>% pull(sd) %>% round_x(., 2)`; increase in log-odds: $\beta$ = `r round_x(s3_gam_ptab["factorsoc_cog", "Estimate"], 2)`, $p$ `r ifelse(s3_gam_ptab["factorsoc_cog", "Pr(>|z|)"] < 0.001, " < 0.001", paste(" = ", round_x(s3_gam_ptab["factorsoc_cog", "Pr(>|z|)"], 3)))`) and _negative affect_ (median score for newborns: `r s3_cap_rat_newborns_ss %>% filter(factor == "Negative affect") %>% pull(median) %>% round_x(., 2)` out of 100 points, $M$ = `r s3_cap_rat_newborns_ss %>% filter(factor == "Negative affect") %>% pull(mean) %>% round_x(., 2)`, $sd$ = `r s3_cap_rat_newborns_ss %>% filter(factor == "Negative affect") %>% pull(sd) %>% round_x(., 2)`; increase in log-odds: $\beta$ = `r round_x(s3_gam_ptab["factorneg_cog", "Estimate"], 2)`, $p$ `r ifelse(s3_gam_ptab["factorneg_cog", "Pr(>|z|)"] < 0.001, " < 0.001", paste(" = ", round_x(s3_gam_ptab["factorneg_cog", "Pr(>|z|)"], 3)))`); in Study 3, as in Study 2, participants reported substantial increases in the domain that we have labeled _social connection_ ($\chi^2$ = `r s3_gam_parameters %>% filter(Parameter == "s(target_year):factorSocial connection") %>% pull(z) %>% round_x(., 2)`, $p$ `r ifelse(s3_gam_parameters %>% filter(Parameter == "s(target_year):factorSocial connection") %>% pull(p)  < 0.001, " < 0.001", paste(" = ", round_x(s3_gam_parameters %>% filter(Parameter == "s(target_year):factorSocial connection") %>% pull(p), 3)))`) and relatively small increases over the target age range in the domain of _negative affect_ ($\chi^2$ = `r s3_gam_parameters %>% filter(Parameter == "s(target_year):factorNegative affect") %>% pull(z) %>% round_x(., 2)`, $p$ `r ifelse(s3_gam_parameters %>% filter(Parameter == "s(target_year):factorNegative affect") %>% pull(p)  < 0.001, " < 0.001", paste(" = ", round_x(s3_gam_parameters %>% filter(Parameter == "s(target_year):factorNegative affect") %>% pull(p), 3)))`). Finally, as in Study 2, in the domain of _bodily sensation_ participants attributed nearly maximal capacity to newborns (median score for newborns: `r s3_cap_rat_newborns_ss %>% filter(factor == "Bodily sensation") %>% pull(median) %>% round_x(., 2)` out of 100 points, $M$ = `r s3_cap_rat_newborns_ss %>% filter(factor == "Bodily sensation") %>% pull(mean) %>% round_x(., 2)`, $sd$ = `r s3_cap_rat_newborns_ss %>% filter(factor == "Bodily sensation") %>% pull(sd) %>% round_x(., 2)`; increase in log-odds relative to _Social connection_: $\beta$ = `r round_x(s3_gam_ptab["factorbod_cog", "Estimate"], 2)`, $p$ `r ifelse(s3_gam_ptab["factorbod_cog", "Pr(>|z|)"] < 0.001, " < 0.001", paste(" = ", round_x(s3_gam_ptab["factorbod_cog", "Pr(>|z|)"], 3)))`) and small increases in capacities for _bodily sensation_ over the target age range ($\chi^2$ = `r s3_gam_parameters %>% filter(Parameter == "s(target_year):factorBodily sensation") %>% pull(z) %>% round_x(., 2)`, $p$ `r ifelse(s3_gam_parameters %>% filter(Parameter == "s(target_year):factorBodily sensation") %>% pull(p)  < 0.001, " < 0.001", paste(" = ", round_x(s3_gam_parameters %>% filter(Parameter == "s(target_year):factorBodily sensation") %>% pull(p), 3)))`).

Again, visual inspection of Figure 2, Panel B, makes it clear that the shape of these perceived developmental trajectories varied substantially across domains, echoing quite precisely the results of our post-hoc exploratory analyses in Study 2.

@@

Why do US adults believe certain aspects of human mental life, but not others, to be present at birth? What do US adults perceive to be driving the development of capacities for what we have called _bodily sensation_, _negative affect_, _social connection_, and _cognition and control_? As a first step in what we hope to be a longer line of work addressing such questions, we asked participants to assess the importance of a variety of possible developmental mechanisms for each of the capacities included in this study. We theorized that when people perceive that a capacity is substantially present at birth, they believe that the capacity is innate, and, conversely, that when people perceive that a capacity develops slowly over years, they believe the capacity is in some sense learned. [For in-depth investigations of laypeople's intuitions about the innateness of cognitive abilities, motor abilities, and emotions, see @wang.feigenson2019, @berent.etal2019, @berent.etal2020.]

@@

Meanwhile, the explanation that "the child's brain changes (for example, brain grows bigger, more or fewer connections between neurons)" was a modal choice for `r s3_mech_choice_modes %>% filter(dev_factor == "brain_changes") %>% nrow()` of our 8 capacities, including helplessness (in the domain of _negative affect_) and social learning (_social connection_), emotional control (_cognition and control_) and reasoning about things (_cognition and control_). However, consensus about this selection was much lower (`r (s3_mech_choice_modes %>% filter(dev_factor == "brain_changes") %>% top_n(-1, prop) %>% pull(prop) %>% round(., 2))*100`-`r (s3_mech_choice_modes %>% filter(dev_factor == "brain_changes") %>% top_n(1, prop) %>% pull(prop) %>% round(., 2))*100`% of participants across capacities). Instead, these capacities were characterized by a more diverse selection of mechanisms, including mechanisms that are not as directly tied to biology: For example, when considering what drives the development of a capacity to learn from other people, roughly equal numbers of participants selected brain changes (`r (s3_mech_choice_modes %>% filter(dev_factor == "brain_changes", grepl("learn", capacity)) %>% pull(prop) %>% round(., 2))*100`%), observations of the social world (`r (s3_mech_choice_ss_c %>% filter(grepl("observes people", dev_factor_text), grepl("learn", capacity)) %>% pull(prop) %>% round(., 2))*100`%), and interactions with others (`r (s3_mech_choice_ss_c %>% filter(grepl("interacts", dev_factor_text), grepl("learn", capacity)) %>% pull(prop) %>% round(., 2))*100`%); and when considering the capacity for controlling one's emotions, participants were nearly as likely to select explicit teaching (`r (s3_mech_choice_ss_c %>% filter(grepl("teach", dev_factor_text), grepl("control", capacity)) %>% pull(prop) %>% round(., 2))*100`%) as they were to select brain changes (`r (s3_mech_choice_modes %>% filter(dev_factor == "brain_changes", grepl("control", capacity)) %>% pull(prop) %>% round(., 2))*100`%).

@@

For capacities in the domain that we have called _bodily sensation_ (Figure 3, top row), "the child is biologically 'preprogrammed' to have this ability" stood out as the highest-rated mechanism by far (median score: `r s3_mech_rat_ss_c %>% filter(grepl("Bodily", Domain)) %>% top_n(1, mean) %>% pull(median) %>% as.numeric() %>% round_x(., 2)` on a scale from 0-6, $M$ = `r s3_mech_rat_ss_c %>% filter(grepl("Bodily", Domain)) %>% top_n(1, mean) %>% pull(mean) %>% as.numeric() %>% round_x(., 2)`, sd = `r s3_mech_rat_ss_c %>% filter(grepl("Bodily", Domain)) %>% top_n(1, mean) %>% pull(sd) %>% as.numeric() %>% round_x(., 2)`; comparison to grand mean: $\beta$ = `r s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("prep", Parameter)) %>% pull(Coefficient) %>% as.numeric() %>% round_x(., 2)`, 95% CI: [`r s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("prep", Parameter)) %>% data.frame() %>% select(CI_low, CI_high) %>% round_x(., 2) %>% unite(CI95, c(CI_low, CI_high), sep = ", ") %>% as.character()`], $p$ `r ifelse(s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("prep", Parameter)) %>% pull(p) < 0.001, " < 0.001", paste(" = ", round_x(s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("prep", Parameter)) %>% pull(p), 3)))`), followed by "the child's brain changes (for example, brain grows bigger, more or fewer connections between neurons)" (median score: `r s3_mech_rat_ss_c %>% filter(grepl("Bodily", Domain)) %>% top_n(2, mean) %>% top_n(-1, mean) %>% pull(median) %>% as.numeric() %>% round_x(., 2)`, $M$ = `r s3_mech_rat_ss_c %>% filter(grepl("Bodily", Domain)) %>% top_n(2, mean) %>% top_n(-1, mean) %>% pull(mean) %>% as.numeric() %>% round_x(., 2)`, sd = `r s3_mech_rat_ss_c %>% filter(grepl("Bodily", Domain)) %>% top_n(2, mean) %>% top_n(-1, mean) %>% pull(sd) %>% as.numeric() %>% round_x(., 2)`; comparison to grand mean: $\beta$ = `r s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("brai", Parameter)) %>% pull(Coefficient) %>% as.numeric() %>% round_x(., 2)` [`r s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("brai", Parameter)) %>% data.frame() %>% select(CI_low, CI_high) %>% round_x(., 2) %>% unite(CI95, c(CI_low, CI_high), sep = ", ") %>% as.character()`], $p$ `r ifelse(s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("brai", Parameter)) %>% pull(p) < 0.001, " < 0.001", paste(" = ", round_x(s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("brai", Parameter)) %>% pull(p), 3)))`), and "the child's body grows and matures (for example, muscles get stronger, child gets taller)" (median score: `r s3_mech_rat_ss_c %>% filter(grepl("Bodily", Domain)) %>% top_n(3, mean) %>% top_n(-1, mean) %>% pull(median) %>% as.numeric() %>% round_x(., 2)`, $M$ = `r s3_mech_rat_ss_c %>% filter(grepl("Bodily", Domain)) %>% top_n(3, mean) %>% top_n(-1, mean) %>% pull(mean) %>% as.numeric() %>% round_x(., 2)`, sd = `r s3_mech_rat_ss_c %>% filter(grepl("Bodily", Domain)) %>% top_n(3, mean) %>% top_n(-1, mean) %>% pull(sd) %>% as.numeric() %>% round_x(., 2)`; comparison to grand mean: $\beta$ = `r s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("body", Parameter)) %>% pull(Coefficient) %>% as.numeric() %>% round_x(., 2)` [`r s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("body", Parameter)) %>% data.frame() %>% select(CI_low, CI_high) %>% round_x(., 2) %>% unite(CI95, c(CI_low, CI_high), sep = ", ") %>% as.character()`], $p$ `r ifelse(s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("body", Parameter)) %>% pull(p) < 0.001, " < 0.001", paste(" = ", round_x(s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("body", Parameter)) %>% pull(p), 3)))`). The remaining mechanisms were rated as quite unimportant for capacities in this domain (all medians $\leq$ `r s3_mech_rat_ss_c %>% filter(grepl("Bodily", Domain)) %>% top_n(-7, median) %>% top_n(1, median) %>% distinct(median) %>% pull(median) %>% as.numeric() %>% round_x(., 2)`; all $M$ $\leq$ `r s3_mech_rat_ss_c %>% filter(grepl("Bodily", Domain)) %>% top_n(-7, mean) %>% top_n(1, mean) %>% distinct(mean) %>% pull(mean) %>% as.numeric() %>% round_x(., 2)`), with particularly low ratings for "people explicitly teach the child how to do this" (median score: `r s3_mech_rat_ss_c %>% filter(grepl("Bodily", Domain)) %>% top_n(-1, mean) %>% pull(median) %>% as.numeric() %>% round_x(., 2)`, $M$ = `r s3_mech_rat_ss_c %>% filter(grepl("Bodily", Domain)) %>% top_n(-1, mean) %>% pull(mean) %>% as.numeric() %>% round_x(., 2)`, sd = `r s3_mech_rat_ss_c %>% filter(grepl("Bodily", Domain)) %>% top_n(-1, mean) %>% pull(sd) %>% as.numeric() %>% round_x(., 2)`; comparison to grand mean: $\beta$ = `r s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("teac", Parameter)) %>% pull(Coefficient) %>% as.numeric() %>% round_x(., 2)` [`r s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("teac", Parameter)) %>% data.frame() %>% select(CI_low, CI_high) %>% round_x(., 2) %>% unite(CI95, c(CI_low, CI_high), sep = ", ") %>% as.character()`], $p$ `r ifelse(s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("teac", Parameter)) %>% pull(p) < 0.001, " < 0.001", paste(" = ", round_x(s3_lmer_bod_tab %>% filter(Effects == "fixed", grepl("teac", Parameter)) %>% pull(p), 3)))`).
